---
title: "Quantium Virtual Internship - Retail Strategy and Analytics - Task 1"
author: "Eda AYDIN"
output: pdf_document
---

## Load required libraries

```{r}
library(readxl)
library(dplyr)
library(data.table)
```

## Define File Path and Load Data

```{r}
filePath <- ""  # Set working directory
# read excel file
transactions <- readxl::read_excel(paste0(filePath, "QVI_transaction_data.xlsx"))
customers <- fread(paste0(filePath, "QVI_purchase_behaviour.csv"))
```

## Display the data

```{r}
head(transactions)
```

```{r}
head(customers)
```

## Summary Statistics

```{r}
summary(transactions)
```

```{r}
summary(customers)
```

```{r}
# number of rows
nrow(transactions)
```

```{r}
nrow(customers)
```

## Variables Description

The transaction data contains the following variables:

-   **DATE**: Date of purchase
-   **STORE_NBR**: Store number
-   **LYLTY_CARD_NBR**: Customer loyalty card number
-   **TXN_ID**: Transaction ID
-   **PROD_NBR**: Product number
-   **PROD_NAME**: Product name
-   **PROD_QTY**: Quantity of product purchased
-   **TOT_SALES**: Total sales (\$)

The customer data contains the following variables:

-   **LYLTY_CARD_NBR**: Customer loyalty card number
-   **LIFESTAGE**: Customer lifestage
-   **PREMIUM_CUSTOMER**: Customer premium status

## Data Cleaning

### Missing Values

```{r}
colSums(is.na(transactions))
```

```{r}
colSums(is.na(customers))
```

### Fix the type

```{r}
transactions$DATE <- as.Date(transactions$DATE, origin="1899-12-30")
head(transactions)
```

### Outlier Detection

```{r}
hist(transactions$TOT_SALES, main="Histogram of Total Sales", xlab="Total Sales ($)")
```

```{r}
# remove outliers
q1 <- quantile(transactions$TOT_SALES, 0.25, na.rm = TRUE)  # First quartile
q3 <- quantile(transactions$TOT_SALES, 0.75, na.rm = TRUE)  # First quartile
IQR <- q3 - q1

lower_bound <- q1 - 1.5 * IQR
upper_bound <- q3 + 1.5 * IQR

transactions <- transactions[transactions$TOT_SALES <= upper_bound, ]
```

```{r}
hist(transactions$TOT_SALES, main="Histogram of Total Sales", xlab="Total Sales ($)")
```

# Merge data

```{r}
merged_data <- merge(transactions, customers, by="LYLTY_CARD_NBR")
head(merged_data)
```

# Exploratory Data Analysis

```{r}
# Analyze the young singles/couples
young_singles_couples <- merged_data %>% filter(LIFESTAGE %in% c("YOUNG SINGLES/COUPLES"))
head(young_singles_couples)
```

```{r}
summary(young_singles_couples)
str(young_singles_couples)
```

```{r}
sum(young_singles_couples$TOT_SALES, na.rm = TRUE)
```

```{r}
hist(young_singles_couples$TOT_SALES, main="Histogram of Total Sales for Young Singles/Couples", xlab="Total Sales ($)")
```

```{r}
# Summarize total sales by product name
product_sales <- young_singles_couples %>%
  group_by(PROD_NAME) %>%
  summarise(Total_Sales = sum(TOT_SALES, na.rm = TRUE)) %>%
  arrange(desc(Total_Sales))  # Optional: Sort by total sales in descending order
head(product_sales)
```

```{r}
# Summarize total sales by premium status
premium_status_sales <- young_singles_couples %>%
        group_by(PREMIUM_CUSTOMER) %>%
        summarize(Total_sales = sum(TOT_SALES, na.rm = TRUE)) %>%
        arrange(desc(Total_sales))  # Optional: Sort by total sales in descending order
premium_status_sales

# Barplot for total sales by premium status
barplot(premium_status_sales$Total_sales, names.arg = premium_status_sales$PREMIUM_CUSTOMER, main = "Total Sales by Premium Status for Young Singles/Couples", xlab = "Premium Status", ylab = "Total Sales ($)", col = "skyblue")
```

```{r}
# Sample Mainstream Customers
mainstream_customers <- young_singles_couples[young_singles_couples$PREMIUM_CUSTOMER == "Mainstream", ]

# Summarize total sales by product name for mainstream customers
product_sales <- mainstream_customers %>%
  group_by(PROD_NAME) %>%
  summarise(Total_Sales = sum(TOT_SALES, na.rm = TRUE)) %>%
  arrange(desc(Total_Sales))  # Optional: Sort by total sales in descending order
head(product_sales)
```

```{r}
# Sample Mainstream Customers
budget_customers <- young_singles_couples[young_singles_couples$PREMIUM_CUSTOMER == "Budget", ]

# Summarize total sales by product name for mainstream customers
product_sales <- budget_customers %>%
  group_by(PROD_NAME) %>%
  summarise(Total_Sales = sum(TOT_SALES, na.rm = TRUE)) %>%
  arrange(desc(Total_Sales))  # Optional: Sort by total sales in descending order
head(product_sales)
```

```{r}
# Sample Mainstream Customers
premium_customers <- young_singles_couples[young_singles_couples$PREMIUM_CUSTOMER == "Premium", ]

# Summarize total sales by product name for mainstream customers
product_sales <- premium_customers %>%
  group_by(PROD_NAME) %>%
  summarise(Total_Sales = sum(TOT_SALES, na.rm = TRUE)) %>%
  arrange(desc(Total_Sales))  # Optional: Sort by total sales in descending order
head(product_sales)
```

```{r}
library(ggplot2)
library(data.table)

#### Aggregate Sales Data
sales_summary <- merged_data %>%
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>%
  summarise(Total_Sales = sum(TOT_SALES, na.rm = TRUE))

#### Plot Total Sales by Customer Segment
ggplot(sales_summary, aes(x = LIFESTAGE, y = Total_Sales, fill = PREMIUM_CUSTOMER)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_col(position = "dodge") +
  labs(title = "Total Sales by Customer Segment", x = "Lifestage", y = "Total Sales")
```

```{r}
t_test <- t.test(
  merged_data$TOT_SALES[merged_data$PREMIUM_CUSTOMER == "Mainstream" & merged_data$LIFESTAGE == "YOUNG SINGLES/COUPLES"],
  merged_data$TOT_SALES[merged_data$PREMIUM_CUSTOMER == "Premium" & merged_data$LIFESTAGE == "YOUNG SINGLES/COUPLES"]
)
  
print(t_test)
```
