---
title: "Quantium Virtual Internship - Retail Strategy and Analytics - Task 2"
mainfont: Times
monofont: Helvetica
output:
  pdf_document:
    df_print: default
    highlight: tango
    keep_tex: yes
    latex_engine: xelatex
header-includes:
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(linewidth=80)
```
```{r knitr line wrap setup, include=FALSE}
library(knitr)
hook_output = knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  if (!is.null(n <- options$linewidth)) {
    x <- knitr:::split_lines(x)
    if (any(nchar(x) > n))
      x <- strwrap(x, width = n)
    x <- paste(x, collapse = "\n")
  }
  hook_output(x, options)
})
```
## Load required libraries and datasets
Note that you will need to install these libraries if you have never used these
before.
```{r 0. Load libraries, include = FALSE}
library(data.table)
library(ggplot2)
library(tidyr)
```
#### Point the filePath to where you have downloaded the datasets to and
#### assign the data files to data.tables
```{r 1. Read in data from previous module}
file_path <- ""
data <- fread(paste0(filePath, "QVI_data.csv"))
#### Set themes for plots
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
```

## Select control stores
The client has selected store numbers 77, 86 and 88 as trial stores and want
control stores to be established stores that are operational for the entire
observation period.
We would want to match trial stores to control stores that are similar to the trial
store prior to the trial period of Feb 2019 in terms of :
- Monthly overall sales revenue
- Monthly number of customers
- Monthly number of transactions per customer
Let's first create the metrics of interest and filter to stores that are present
throughout the pre-trial period.
```{r Select control stores}
#### Calculate these measures over time for each store
# Add a new column YEARMONTH in the format yyyymm
data[, YEARMONTH := year(DATE)*100 + month(DATE)]

measure_over_time <- data[, .(
  totSales = sum(TOT_SALES),
  nCustomers = uniqueN(LYLTY_CARD_NBR),
  nTxnPerCust = uniqueN(TXN_ID)/uniqueN(LYLTY_CARD_NBR),
  nChipsPerTxn = sum(PROD_QTY)/uniqueN(TXN_ID),
  avgPricePerUnit = sum(TOT_SALES)/sum(PROD_QTY)
), by = .(STORE_NBR, YEARMONTH)]

# Order the results by store number and yearmonth for readability
measure_over_time <- measure_over_time[order(STORE_NBR, YEARMONTH)]

# Find stores with full observations over the pre-trial period
# Here, we assume that the pre-trial period has 12 months. 
# Adjust this if needed.
full_obs_stores <- measure_over_time[YEARMONTH < 201902, 
                                     .N, by = STORE_NBR][N == 7]$STORE_NBR

# Filter for pre-trial measures and stores with full observation periods
pretrial_measures <- measure_over_time[YEARMONTH < 201902 & 
                                        STORE_NBR %in% full_obs_stores,]
```
Now we need to work out a way of ranking how similar each potential control store
is to the trial store. We can calculate how correlated the performance of each
store is to the trial store.
Let's write a function for this so that we don't have to calculate this for each
trial store and control store pair.
```{r Create function to calculate correlation}
data[, DATE := as.IDate(DATE, format = "%Y-%m-%d")]
data[, YEARMONTH := year(DATE) * 100 + month(DATE)]

# Define the metrics calculations for each store and month
measure_over_time <- data[, .(
  totSales = sum(TOT_SALES),
  nCustomers = uniqueN(LYLTY_CARD_NBR),
  nTxnPerCust = .N / uniqueN(LYLTY_CARD_NBR),
  nChipsPerTxn = sum(PROD_QTY) / .N,
  avgPricePerUnit = sum(TOT_SALES) / sum(PROD_QTY)
), by = .(STORE_NBR, YEARMONTH)
][order(STORE_NBR, YEARMONTH)]

# Filter for the pre-trial period and stores with full observation periods
full_obs_stores <- measure_over_time[YEARMONTH < 201902,
                                     .N, by = STORE_NBR][N == 7, STORE_NBR]
pretrial_measures <- measure_over_time[YEARMONTH < 201902 &
                                         STORE_NBR %in% full_obs_stores, ]

# Function to calculate correlation
calculate_correlation <- function(input_table, metric_col, store_comparison) {
  store_numbers <- unique(
    input_table$STORE_NBR[input_table$STORE_NBR != store_comparison]
  )
  list_to_bind <- list()

  for (i in store_numbers) {
    if (i != store_comparison) {
      # Using .SD to specify columns dynamically
      comparison_values <- input_table[STORE_NBR == store_comparison, .SD[[metric_col]], .SDcols = metric_col]
      comparison_i_values <- input_table[STORE_NBR == i, .SD[[metric_col]], .SDcols = metric_col]
      calc_measure <- cor(comparison_values, comparison_i_values, use = "complete.obs")
      calculated_measure <- data.table(
        Store1 = store_comparison,
        Store2 = i,
        corr_measure = calc_measure
      )
      list_to_bind[[length(list_to_bind) + 1]] <- calculated_measure
    }
  }

  calc_corr_table <- rbindlist(list_to_bind, fill = TRUE)
  return(calc_corr_table[order(-corr_measure)])
}

# Now you can call this function with your pretrial_measures and metrics
# For example, to compare store 77:
corr_sales <- calculate_correlation(pretrial_measures, "totSales", 77)
corr_customers <- calculate_correlation(pretrial_measures, "nCustomers", 77)

# Display the top correlated stores for sales and customers
print(corr_sales)
print(corr_customers)
```
Apart from correlation, we can also calculate a standardised metric based on the
absolute difference between the trial store's performance and each control store's
performance.
Let's write a function for this.
```{r Create function to calculate magnitude distance}
calculateMagnitudeDistance <- function(inputTable, metricCol, storeComparison) {
  storeNumbers <- unique(inputTable[STORE_NBR != storeComparison, STORE_NBR])
  calcDistTable <- data.table(Store1 = numeric(), Store2 = numeric(), YEARMONTH = numeric(), measure = numeric())

  for (i in storeNumbers) {
    if (i != storeComparison) {
      calculatedMeasure <- data.table(
        Store1 = storeComparison,
        Store2 = i,
        YEARMONTH = inputTable[STORE_NBR == storeComparison, YEARMONTH],
        measure = abs(inputTable[STORE_NBR == storeComparison, ..metricCol] - 
                      inputTable[STORE_NBR == i, ..metricCol])
      )
      calcDistTable <- rbind(calcDistTable, calculatedMeasure)
    }
  }
  minMaxDist <- calcDistTable[, .(minDist = min(measure), maxDist = max(measure)), by = .(Store1, YEARMONTH)]
  distTable <- merge(calcDistTable, minMaxDist, by = c("Store1", "YEARMONTH"))
  distTable[, magnitudeMeasure := 1 - (measure - minDist) / (maxDist - minDist)]
  finalDistTable <- distTable[, .(mag_measure = mean(magnitudeMeasure, na.rm = TRUE)), by = .(Store1, Store2)]
  
  return(finalDistTable)
}
```
Now let's use the functions to find the control stores! We'll select control stores
based on how similar monthly total sales in dollar amounts and monthly number of
customers are to the trial stores. So we will need to use our functions to get four
scores, two for each of total sales and total customers.

```{r Use functions to calculate metrics}
trial_store <- 77 

# Calculating correlation for sales and number of customers
corr_nSales <- calculate_correlation(pretrial_measures, "totSales", trial_store)
corr_nCustomers <- calculate_correlation(pretrial_measures, "nCustomers", trial_store)

# Calculating magnitude distance for sales and number of customers
magnitude_nSales <- calculateMagnitudeDistance(pretrial_measures, "totSales", trial_store)
magnitude_nCustomers <- calculateMagnitudeDistance(pretrial_measures, "nCustomers", trial_store)
```

We'll need to combine the all the scores calculated using our function to create a
composite score to rank on.
Let's take a simple average of the correlation and magnitude scores for each
driver. Note that if we consider it more important for the trend of the drivers to
be similar, we can increase the weight of the correlation score (a simple average
gives a weight of 0.5 to the corr_weight) or if we consider the absolute size of
the drivers to be more important, we can lower the weight of the correlation score.
```{r}
score_nSales <- merge(corr_nSales, magnitude_nSales, by = c("Store1", "Store2"))
score_nCustomers <- merge(corr_nCustomers, magnitude_nCustomers, by = c("Store1", "Store2"))

# Calculating the combined score for sales and customers
score_nSales[, scoreNSales := (corr_measure + mag_measure) / 2]
score_nCustomers[, scoreNCust := (corr_measure + mag_measure) / 2]

# Merging the combined scores for sales and customers into a final score
score_Control <- merge(score_nSales[, .(Store1, Store2, scoreNSales)], score_nCustomers[, .(Store1, Store2, scoreNCust)], by = c("Store1", "Store2"))
score_Control[, finalControlScore := (scoreNSales + scoreNCust) / 2]

# Selecting the control store based on the highest final combined score
bestControlStore <- score_Control[order(-finalControlScore)][2]  # Assuming the first row is the trial store itself

# Display the best control store
print(bestControlStore)
```
The store with the highest score is then selected as the control store since it is
most similar to the trial store.
```{r}
bestControlStoreFor77 <- combinedCorrScores[Store1 == 77][1, ]
control_store <- bestControlStoreFor77$Store2
```
Now that we have found a control store, let's check visually if the drivers are
indeed similar in the period before the trial.
We'll look at total sales first.
```{r}
#### Visual checks on trends based on the drivers
trial_store <- 77
control_store <- bestControlStoreFor77$Store2

measure_over_time[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
                                       ifelse(STORE_NBR == control_store, "Control", "Other stores"))]

measure_over_timeSales <- measure_over_time[YEARMONTH < 201903, .(totSales = sum(totSales)), by = .(Store_type, YEARMONTH)]
measure_over_timeSales[, TransactionMonth := as.Date(paste0(YEARMONTH %/% 100, "-", YEARMONTH %% 100, "-01"))]

ggplot(measure_over_timeSales, aes(x = TransactionMonth, y = totSales, color = Store_type)) +
  geom_line() +
  labs(title = "Total sales by month", x = "Month of operation", y = "Total sales") +
  theme_minimal() +
  scale_color_manual(values = c("Trial" = "blue", "Control" = "red", "Other stores" = "grey50")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Next, number of customers.
```{r}
measure_over_timeCusts <- measure_over_time[YEARMONTH < 201903, .(nCustomers), by = .(Store_type, YEARMONTH)]
measure_over_timeCusts[, TransactionMonth := as.Date(paste0(substr(YEARMONTH, 1, 4), "-", substr(YEARMONTH, 5, 6), "-01"))]

ggplot(measure_over_timeCusts, aes(x = TransactionMonth, y = nCustomers, color = Store_type)) +
  geom_line() +
  labs(title = "Number of Customers by Month", x = "Month of Operation", y = "Number of Customers") +
  theme_minimal() +
  scale_color_manual(values = c("Trial" = "blue", "Control" = "red", "Other stores" = "grey50")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
## Assessment of trial
The trial period goes from the start of February 2019 to April 2019. We now want to
see if there has been an uplift in overall chip sales.
We'll start with scaling the control store's sales to a level similar to control
for any differences between the two stores outside of the trial period.
```{r Comparison of results during trial}
#### Scale pre-trial control sales to match pre-trial trial store sales
scalingFactorForControlSales <- pretrial_measures[STORE_NBR == trial_store &
YEARMONTH < 201902, sum(totSales)]/pretrial_measures[STORE_NBR == control_store &
YEARMONTH < 201902, sum(totSales)]
#### Apply the scaling factor
measure_over_timeSales <- measure_over_time
scaledControlSales <- measure_over_timeSales[STORE_NBR == control_store, ][ ,
controlSales := totSales * scalingFactorForControlSales]
```
Now that we have comparable sales figures for the control store, we can calculate
the percentage difference between the scaled control sales and the trial store's
sales during the trial period.
```{r}
trialSales <- measure_over_timeSales[STORE_NBR == trial_store & YEARMONTH %in% 201902:201904]
scaledControlSalesTrialPeriod <- scaledControlSales[YEARMONTH %in% 201902:201904]
percentageDiff <- merge(trialSales, scaledControlSalesTrialPeriod, by = "YEARMONTH")
percentageDiff[, percentageDiff := abs((totSales.x - controlSales) / controlSales) * 100]
percentageDiff <- percentageDiff[, .(YEARMONTH, TrialSales = totSales.x, ScaledControlSales = controlSales, PercentageDifference = percentageDiff)]
```
Let's see if the difference is significant!
```{r}
stdDev <- sd(percentageDiff[YEARMONTH < 201902, "PercentageDifference", with = FALSE]$PercentageDifference)
meanPreTrial <- mean(percentageDiff[YEARMONTH < 201902, "PercentageDifference", with = FALSE]$PercentageDifference)
percentageDiff[, tValue := (PercentageDifference - meanPreTrial) / stdDevPreTrial]

degreesOfFreedom <- 7 

criticalValue <- qt(0.95, df = degreesOfFreedom)
percentageDiff[, Significance := ifelse(abs(tValue) > criticalValue, "Significant", "Not Significant")]
percentageDiff[, TransactionMonth := as.Date(paste0(substr(YEARMONTH, 1, 4), "-", substr(YEARMONTH, 5, 6), "-01"))]
```
We can observe that the t-value is much larger than the 95th percentile value of
the t-distribution for March and April - i.e. the increase in sales in the trial
store in March and April is statistically greater than in the control store.
Let's create a more visual version of this by plotting the sales of the control
store, the sales of the trial stores and the 95th percentile value of sales of the
control store.
```{r, fig.align = "Center"}
measure_over_timeSales[, TransactionMonth := as.IDate(paste(YEARMONTH %/% 100, YEARMONTH %% 100, "01", sep="-"), "%Y-%m-%d")]
measure_over_timeSales[, Store_type := ifelse(STORE_NBR == trial_store, "Trial", 
                                            ifelse(STORE_NBR == control_store, "Control", "Other"))]
trialAndControlSales <- measure_over_timeSales[Store_type %in% c("Trial", "Control"),]
trialAndControlSales[, totSales95 := ifelse(Store_type == "Control", totSales * (1 + stdDev * 2), totSales)]

trialAndControlSales[, totSales5 := ifelse(Store_type == "Control", totSales * (1 - stdDev * 2), totSales)]
trialAssessment <- melt(trialAndControlSales, id.vars = c("TransactionMonth", "Store_type"), 
                        measure.vars = c("totSales", "totSales95", "totSales5"),
                        variable.name = "Measure", value.name = "Sales")
trialAssessment[, Store_type := ifelse(Measure == "totSales95", paste0(Store_type, " - 95th percentile"),
                                        ifelse(Measure == "totSales5", paste0(Store_type, " - 5th percentile"), Store_type))]

ggplot(data = trialAssessment, aes(x = TransactionMonth, y = Sales, color = Store_type)) +
  geom_line() +
  geom_rect(
    aes(xmin = as.Date("2019-02-01"), xmax = as.Date("2019-04-30"), ymin = -Inf, ymax = Inf),
    fill = "grey80", alpha = 0.2, inherit.aes = FALSE
  ) +
  scale_color_manual(values = c("Trial" = "blue", "Control" = "red", "Control - 95th percentile" = "red", "Control - 5th percentile" = "red")) +
  labs(title = "Total Sales by Month", x = "Month", y = "Total Sales", color = "Store Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
The results show that the trial in store 77 is significantly different to its
control store in the trial period as the trial store performance lies outside the
5% to 95% confidence interval of the control store in two of the three trial
months.
Let's have a look at assessing this for number of customers as well.
```{r }
measure_over_time <- data[, .(
  nCustomers = uniqueN(LYLTY_CARD_NBR)
), by = .(STORE_NBR, YEARMONTH)]

scalingFactorForControlCust <- pretrial_measures[STORE_NBR == trial_store & YEARMONTH < 201902, sum(nCustomers)] /
                              pretrial_measures[STORE_NBR == control_store & YEARMONTH < 201902, sum(nCustomers)]

measure_over_timeCusts <- measure_over_time[, .(
  ScaledControlCustomers = as.numeric(ifelse(STORE_NBR == control_store, nCustomers * scalingFactorForControlCust, nCustomers))
), by = .(STORE_NBR, YEARMONTH)]
trialPeriodCusts <- measure_over_timeCusts[STORE_NBR %in% c(trial_store, control_store) & YEARMONTH %in% c(201902, 201903, 201904)]
trialPeriodCusts <- merge(trialPeriodCusts, measure_over_time[STORE_NBR == trial_store, .(STORE_NBR, YEARMONTH, nCustomers)], by = c("STORE_NBR", "YEARMONTH"), all.x = TRUE)
trialPeriodCusts[, PercentageDifference := ifelse(STORE_NBR == trial_store, 
                                                  abs((ScaledControlCustomers - nCustomers) / ScaledControlCustomers) * 100, NA)]

ggplot(trialPeriodCusts, aes(x = YEARMONTH, y = ScaledControlCustomers, group = STORE_NBR, color = factor(STORE_NBR))) +
  geom_line() +
  geom_point() +
  geom_line(data = subset(trialPeriodCusts, STORE_NBR == trial_store), aes(y = nCustomers)) +
  geom_point(data = subset(trialPeriodCusts, STORE_NBR == trial_store), aes(y = nCustomers)) +
  scale_color_manual(values = c(trial_store = "blue", control_store = "red")) +
  labs(title = "Comparison of Customer Counts", x = "YearMonth", y = "Number of Customers", color = "Store Number") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Let's again see if the difference is significant visually!
```{r , fig.align = "Center"}
measure_over_time[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
                                       ifelse(STORE_NBR == control_store, "Control", "Other"))]

# Calculate scaling factor for control customers
scalingFactorForControlCust <- sum(measure_over_time[STORE_NBR == trial_store & YEARMONTH < 201902, nCustomers]) /
                               sum(measure_over_time[STORE_NBR == control_store & YEARMONTH < 201902, nCustomers])

# Apply the scaling factor
measure_over_time[, ScaledControlCustomers := ifelse(STORE_NBR == control_store, round(nCustomers * scalingFactorForControlCust), nCustomers)]

# Now, proceed with your original plan of plotting, but now referencing 'ScaledControlCustomers' where necessary
pastCustomers <- measure_over_time[, .(nCusts = mean(ScaledControlCustomers, na.rm = TRUE)), by = .(YEARMONTH, Store_type)]
pastCustomers_Controls95 <- pastCustomers[Store_type == "Control", ][, nCusts95 := nCusts * (1 + stdDev * 2)]
pastCustomers_Controls5 <- pastCustomers[Store_type == "Control", ][, nCusts5 := nCusts * (1 - stdDev * 2)]

# Combine the above data for plotting
trialAssessmentCustomers <- rbind(
  pastCustomers, 
  setNames(pastCustomers_Controls95[, .(YEARMONTH, Store_type, nCusts = nCusts95)], c("YEARMONTH", "Store_type", "nCusts")),
  setNames(pastCustomers_Controls5[, .(YEARMONTH, Store_type, nCusts = nCusts5)], c("YEARMONTH", "Store_type", "nCusts"))
)

# Proceed with ggplot as before
ggplot(trialAssessmentCustomers, aes(x = YEARMONTH, y = nCusts, color = Store_type)) +
  geom_line() +
  geom_point() +
  geom_rect(data = trialAssessmentCustomers[YEARMONTH %in% c(201902, 201903, 201904),],
            aes(xmin = min(YEARMONTH), xmax = max(YEARMONTH), ymin = 0, ymax = Inf, fill = Store_type),
            alpha = 0.2, inherit.aes = FALSE, show.legend = FALSE) +
  scale_color_manual(values = c("Trial" = "blue", "Control" = "red", "Other stores" = "grey")) +
  labs(title = "Number of Customers: Trial vs Control Store", x = "Year-Month", y = "Number of Customers") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Let's repeat finding the control store and assessing the impact of the trial for
each of the other two trial stores.
## Trial store 86
```{r}
trial_store <- 86

corr_sales <- calculate_correlation(pretrial_measures, "totSales", trial_store)
corr_customers <- calculate_correlation(pretrial_measures, "nCustomers", trial_store)

magnitude_sales <- calculateMagnitudeDistance(pretrial_measures, "totSales", trial_store)
magnitude_customers <- calculateMagnitudeDistance(pretrial_measures, "nCustomers", trial_store)

score_nSales <- merge(corr_nSales, magnitude_nSales, by = c("Store1", "Store2"))
score_nSales[, scoreNSales := (corr_measure + mag_measure) / 2]

score_nCustomers <- merge(corr_nCustomers, magnitude_nCustomers, by = c("Store1", "Store2"))
score_nCustomers[, scoreNCust := (corr_measure + mag_measure) / 2]

score_Control <- merge(score_nSales[, .(Store1, Store2, scoreNSales)], 
                       score_nCustomers[, .(Store1, Store2, scoreNCust)], 
                       by = c("Store1", "Store2"))
score_Control[, finalControlScore := (scoreNSales + scoreNCust) / 2]

control_store <- score_Control[Store1 == trial_store & Store2 != trial_store][order(-finalControlScore)][1, .(Store2, finalControlScore)]
```
Looks like store 155 will be a control store for trial store 86.
Again, lets check visually if the drivers are indeed similar in the period before
the trial.
Well look at total sales first.
```{r}
measure_over_timeSales <- measure_over_time[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
                                            ifelse(STORE_NBR == control_store, "Control", "Other stores"))]

pastSales <- measure_over_timeSales[YEARMONTH < 201903, .(totSales = sum(totSales)), by = .(TransactionMonth = YEARMONTH, Store_type)]

ggplot(data = pastSales, aes(x = as.Date(paste0(TransactionMonth, "01"), format = "%Y%m%d"), y = totSales, color = Store_type)) +
  geom_line() +
  labs(title = "Total Sales by Month", x = "Month", y = "Total Sales", color = "Store Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Great, sales are trending in a similar way.
Next, number of customers.
```{r}
# Assign a store type label to each store number for easy identification in the plots
measure_over_timeSales[, Store_type := ifelse(STORE_NBR == 86, "Trial",
                                 ifelse(STORE_NBR == 155, "Control", "Other"))]

pastSales <- measure_over_timeSales[YEARMONTH < 201902, 
                                  .(totSales = sum(totSales)), by = .(YEARMONTH, Store_type)]

pastSales[, TransactionMonth := as.Date(paste0(substr(YEARMONTH, 1, 4), "-", substr(YEARMONTH, 5, 6), "-01"))]

ggplot(data = pastSales, aes(x = TransactionMonth, y = totSales, group = Store_type, color = Store_type)) +
  geom_line() +
  labs(x = "Month", y = "Total Sales", title = "Total Monthly Sales by Store Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
  scale_color_manual(values = c("Trial" = "blue", "Control" = "green", "Other" = "grey"))
```
Good, the trend in number of customers is also similar.
Let's now assess the impact of the trial on sales.
```{r, fig.align = "Center"}
#### Scale pre-trial control sales to match pre-trial trial store sales
scalingFactorForControlSales <- pretrial_measures[STORE_NBR == trial_store & YEARMONTH < 201902, sum(totSales)] /
                                pretrial_measures[STORE_NBR == control_store & YEARMONTH < 201902, sum(totSales)]
measure_over_timeSales[STORE_NBR == control_store, scaledSales := totSales * scalingFactorForControlSales]
percentageDiff <- measure_over_timeSales[STORE_NBR == trial_store & YEARMONTH %in% 201902:201904,
                                       .(TrialSales = sum(totSales), YEARMONTH)][
                   measure_over_timeSales[STORE_NBR == control_store & YEARMONTH %in% 201902:201904],
                   on = "YEARMONTH",
                   .(ControlSales = sum(scaledSales))][, 
                   percentageDiff := abs((TrialSales - ControlSales) / ControlSales)]
stdDev <- measure_over_timeSales[STORE_NBR == control_store & YEARMONTH < 201902, 
                               .(StdDev = sd(scaledSales, na.rm = TRUE))]

degreesOfFreedom <- 7
measure_over_timeSales[STORE_NBR == control_store, `:=` (
  Sales5th = scaledSales - (stdDev * 2),
  Sales95th = scaledSales + (stdDev * 2)
), by = YEARMONTH]
pastSales <- melt(measure_over_timeSales[STORE_NBR %in% c(trial_store, control_store) & YEARMONTH < 201905,
                                       .(TransactionMonth = as.IDate(paste0(YEARMONTH %/% 100, "-", YEARMONTH %% 100, "-01")),
                                         Store_type = ifelse(STORE_NBR == trial_store, "Trial", "Control"),
                                         Sales = totSales,
                                         Sales5th, Sales95th)],
                  id.vars = c("TransactionMonth", "Store_type"), measure.vars = c("Sales", "Sales5th", "Sales95th"))

ggplot(pastSales, aes(x = TransactionMonth, y = value, group = Store_type, color = Store_type)) +
  geom_line() +
  geom_rect(data = subset(pastSales, YEARMONTH >= 201902 & YEARMONTH <= 201904),
            aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = -Inf, ymax = Inf, fill = Store_type),
            alpha = 0.2, inherit.aes = FALSE) +
  scale_color_manual(values = c("Trial" = "blue", "Control" = "darkgreen")) +
  labs(title = "Total Sales by Month", x = "Month", y = "Total Sales", color = "Store Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
The results show that the trial in store 86 is not significantly different to its
control store in the trial period as the trial store performance lies inside the 5%
to 95% confidence interval of the control store in two of the three trial months.
Let's have a look at assessing this for the number of customers as well.
```{r , fig.align = "Center"}
#### This would be a repeat of the steps before for total sales
scalingFactorForControlCust <- pretrial_measures[STORE_NBR == trial_store & YEARMONTH < 201902, sum(nCustomers)] /
                              pretrial_measures[STORE_NBR == control_store & YEARMONTH < 201902, sum(nCustomers)]

measure_over_timeCusts[STORE_NBR == control_store, ScaledControlCustomers := nCustomers * scalingFactorForControlCust]

percentageDiff <- merge(measure_over_timeCusts[STORE_NBR == trial_store & YEARMONTH %in% c(201902, 201903, 201904), .(YEARMONTH, TrialCustomers = nCustomers)],
                        measure_over_timeCusts[STORE_NBR == control_store & YEARMONTH %in% c(201902, 201903, 201904), .(YEARMONTH, ScaledControlCustomers)],
                        by = "YEARMONTH")
percentageDiff[, PercentageDifference := abs(TrialCustomers - ScaledControlCustomers) / ScaledControlCustomers]

stdDev <- sd(percentageDiff[YEARMONTH < 201902, PercentageDifference])
degreesOfFreedom <- 7
measure_over_timeCusts[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
                                             ifelse(STORE_NBR == control_store, "Control", "Other"))]

pastCustomers <- measure_over_timeCusts[Store_type %in% c("Trial", "Control"), .(YEARMONTH, Store_type, nCustomers)]
pastCustomers_Controls95 <- pastCustomers[Store_type == "Control", .(nCustomers95 = nCustomers * (1 + stdDev * 2), YEARMONTH, Store_type)]
pastCustomers_Controls5 <- pastCustomers[Store_type == "Control", .(nCustomers5 = nCustomers * (1 - stdDev * 2), YEARMONTH, Store_type)]

trialAssessment <- merge(pastCustomers, pastCustomers_Controls95, by = c("YEARMONTH", "Store_type"), all = TRUE)
trialAssessment <- merge(trialAssessment, pastCustomers_Controls5, by = c("YEARMONTH", "Store_type"), all = TRUE)

ggplot(data = trialAssessment, aes(x = as.Date(paste0(YEARMONTH, "01"), format = "%Y%m%d"), y = nCustomers, group = Store_type, color = Store_type)) +
  geom_line() +
  geom_ribbon(aes(ymin = nCustomers5, ymax = nCustomers95), data = trialAssessment[Store_type == "Control",], fill = "blue", alpha = 0.2) +
  scale_color_manual(values = c("Trial" = "red", "Control" = "blue")) +
  labs(title = "Number of Customers by Month", x = "Month", y = "Number of Customers") +
  theme_minimal()
```

It looks like the number of customers is significantly higher in all of the three
months. This seems to suggest that the trial had a significant impact on increasing
the number of customers in trial store 86 but as we saw, sales were not
significantly higher. We should check with the Category Manager if there were
special deals in the trial store that were may have resulted in lower prices,
impacting the results.
## Trial store 88
```{r}
trial_store <- 88

corr_nSales <- calculate_correlation(pretrial_measures, 'totSales', trial_store)
corr_nCustomers <- calculate_correlation(pretrial_measures, "nCustomers", trial_store)

magnitude_nSales <- calculateMagnitudeDistance(pretrial_measures, 'totSales', trial_store)
magnitude_nCustomers <- calculateMagnitudeDistance(pretrial_measures, "nCustomers", trial_store)

score_nSales <- merge(corr_nSales, magnitude_nSales, by = c("Store1", "Store2"))
score_nSales[, scoreNSales := (corr_measure + mag_measure) / 2]

score_nCustomers <- merge(corr_nCustomers, magnitude_nCustomers, by = c("Store1", "Store2"))
score_nCustomers[, scoreNCust := (corr_measure + mag_measure) / 2]

score_Control <- merge(score_nSales, score_nCustomers, by = c("Store1", "Store2"))
score_Control[, finalControlScore := (scoreNSales + scoreNCust) / 2]

control_store <- score_Control[Store1 == trial_store, ][order(-finalControlScore)][2, .(Store2)]
```
We've now found store 237 to be a suitable control store for trial store 88.
Again, let's check visually if the drivers are indeed similar in the period before
the trial.
We'll look at total sales first.
```{r}
#### Visual checks on trends based on the drivers
trial_store <- 88

corr_nSales <- calculate_correlation(pretrial_measures, "totSales", trial_store)
corr_nCustomers <- calculate_correlation(pretrial_measures, "nCustomers", trial_store)

magnitude_nSales <- calculateMagnitudeDistance(pretrial_measures, "totSales", trial_store)
magnitude_nCustomers <- calculateMagnitudeDistance(pretrial_measures, "nCustomers", trial_store)

score_nSales <- merge(corr_nSales, magnitude_nSales, by = c("Store1", "Store2"))
score_nSales[, scoreNSales := (corr_measure + mag_measure) / 2]

score_nCustomers <- merge(corr_nCustomers, magnitude_nCustomers, by = c("Store1", "Store2"))
score_nCustomers[, scoreNCust := (corr_measure + mag_measure) / 2]

score_Control <- merge(score_nSales[, .(Store1, Store2, scoreNSales)], score_nCustomers[, .(Store1, Store2, scoreNCust)], by = c("Store1", "Store2"))
score_Control[, finalControlScore := (scoreNSales + scoreNCust) / 2]

control_store <- score_Control[Store1 == trial_store & Store2 != trial_store, .(Store2, finalControlScore)]
control_store <- control_store[which.max(control_store$finalControlScore)]

measure_over_timeSales <- measure_over_timeSales[, Store_type := ifelse(STORE_NBR == trial_store, "Trial", 
                                            ifelse(STORE_NBR == control_store, "Control", "Other stores"))]

pastSales <- measure_over_timeSales[YEARMONTH < 201902, .(totSales = sum(totSales)), by = .(YEARMONTH, Store_type)]

ggplot(data = pastSales, aes(x = as.Date(paste0(YEARMONTH, "01"), format = "%Y%m%d"), y = totSales, color = Store_type)) +
  geom_line() +
  labs(title = "Total Sales by Month", x = "Month", y = "Total Sales", color = "Store Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Great, the trial and control stores have similar total sales.
Next, number of customers.
```{r}
#### Visual checks on trends based on the drivers
# Assuming 'trial_store' is 88 and 'control_store' is 237
trial_store <- 88
control_store <- 237

# Assuming 'measure_over_time' is already defined and contains metrics over time for each store
# Filter for sales data before the trial period for the visual checks
measure_over_timeSales <- measure_over_time[YEARMONTH < 201902,]

pastSales <- measure_over_timeSales[, .(totSales = sum(totSales)), by = .(YEARMONTH, STORE_NBR)]
pastSales$Store_type <- ifelse(pastSales$STORE_NBR == trial_store, "Trial",
                              ifelse(pastSales$STORE_NBR == control_store, "Control", "Other"))
pastSales$TransactionMonth <- as.Date(paste0(YEARMONTH, "01"), "%Y%m%d")

ggplot(data = pastSales, aes(x = TransactionMonth, y = totSales, color = Store_type)) +
  geom_line() +
  labs(title = "Total Sales by Month", x = "Month", y = "Total Sales", color = "Store Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

measure_over_timeCusts <- measure_over_time[YEARMONTH < 201902,]

pastCustomers <- measure_over_timeCusts[, .(nCustomers = uniqueN(LYLTY_CARD_NBR)), by = .(YEARMONTH, STORE_NBR)]
pastCustomers$Store_type <- ifelse(pastCustomers$STORE_NBR == trial_store, "Trial",
                                  ifelse(pastCustomers$STORE_NBR == control_store, "Control", "Other"))
pastCustomers$TransactionMonth <- as.Date(paste0(YEARMONTH, "01"), "%Y%m%d")

ggplot(data = pastCustomers, aes(x = TransactionMonth, y = nCustomers, color = Store_type)) +
  geom_line() +
  labs(title = "Number of Customers by Month", x = "Month", y = "Number of Customers", color = "Store Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Total number of customers of the control and trial stores are also similar.
Let's now assess the impact of the trial on sales.
```{r, fig.align = "Center"}
#### Scale pre-trial control store sales to match pre-trial trial store sales
trial_store <- 88
control_store <- 237 

scalingFactorForControlSales <- sum(pretrial_measures[totSales & STORE_NBR == trial_store]) /
                                sum(pretrial_measures[totSales & STORE_NBR == control_store])

scaledControlSales <- pretrial_measures[STORE_NBR == control_store, 
                                       .(ScaledSales = totSales * scalingFactorForControlSales), 
                                       by = YEARMONTH]

trialPeriodSales <- merge(pretrial_measures[STORE_NBR == trial_store & YEARMONTH %in% c(201902:201904), .(YEARMONTH, TrialSales = totSales)],
                          scaledControlSales[YEARMONTH %in% c(201902:201904), .(YEARMONTH, ScaledSales)],
                          by = "YEARMONTH")

trialPeriodSales[, PercentageDiff := abs(TrialSales - ScaledSales) / ScaledSales * 100]

stdDev <- sd(trialPeriodSales[YEARMONTH < 201902, PercentageDiff])

degreesOfFreedom <- nrow(trialPeriodSales[YEARMONTH < 201902]) - 1

trialPeriodSales[, Sales95th := ScaledSales + (stdDev * 2)]
trialPeriodSales[, Sales5th := ScaledSales - (stdDev * 2)]

ggplot(trialPeriodSales, aes(x = YEARMONTH)) +
  geom_line(aes(y = TrialSales, colour = "Trial Store")) +
  geom_line(aes(y = ScaledSales, colour = "Control Store")) +
  geom_ribbon(aes(ymin = Sales5th, ymax = Sales95th, fill = "Confidence Interval"), alpha = 0.2) +
  labs(title = "Sales Comparison: Trial Store 88 vs Control Store", x = "YearMonth", y = "Sales") +
  scale_fill_manual(values = c("Confidence Interval" = "grey80")) +
  scale_colour_manual(values = c("Trial Store" = "blue", "Control Store" = "red")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
The results show that the trial in store 88 is significantly different to its
control store in the trial period as the trial store performance lies outside of
the 5% to 95% confidence interval of the control store in two of the three trial
months.
Let's have a look at assessing this for number of customers as well.
```{r , fig.align = "Center"}
control_store <- 237

# Calculate scaling factor for customer counts
scalingFactorForControlCust <- sum(pretrial_measures[nCustomers & STORE_NBR == trial_store]) /
                               sum(pretrial_measures[nCustomers & STORE_NBR == control_store])

# Apply scaling factor
scaledControlCustomers <- pretrial_measures[STORE_NBR == control_store,
                                           .(ScaledCustomers = nCustomers * scalingFactorForControlCust), 
                                           by = YEARMONTH]

# Merge trial and scaled control customer counts for the trial period
trialPeriodCustomers <- merge(pretrial_measures[STORE_NBR == trial_store & YEARMONTH %in% c(201902:201904), .(YEARMONTH, TrialCustomers = nCustomers)],
                              scaledControlCustomers[YEARMONTH %in% c(201902:201904), .(YEARMONTH, ScaledCustomers)],
                              by = "YEARMONTH")

# Calculate the percentage difference
trialPeriodCustomers[, PercentageDiff := abs(TrialCustomers - ScaledCustomers) / ScaledCustomers * 100]

# Calculate standard deviation of the percentage difference before the trial
stdDev <- sd(trialPeriodCustomers[YEARMONTH < 201902, PercentageDiff])

# Degrees of freedom
degreesOfFreedom <- nrow(trialPeriodCustomers[YEARMONTH < 201902]) - 1

# Assessing significance visually with 95th and 5th percentile confidence intervals
trialPeriodCustomers[, Customers95th := ScaledCustomers + (stdDev * 2)]
trialPeriodCustomers[, Customers5th := ScaledCustomers - (stdDev * 2)]

# Plotting
ggplot(trialPeriodCustomers, aes(x = YEARMONTH)) +
  geom_line(aes(y = TrialCustomers, colour = "Trial Store")) +
  geom_line(aes(y = ScaledCustomers, colour = "Control Store")) +
  geom_ribbon(aes(ymin = Customers5th, ymax = Customers95th, fill = "Confidence Interval"), alpha = 0.2) +
  labs(title = "Customer Count Comparison: Trial Store 88 vs Control Store", x = "YearMonth", y = "Number of Customers") +
  scale_fill_manual(values = c("Confidence Interval" = "grey80")) +
  scale_colour_manual(values = c("Trial Store" = "blue", "Control Store" = "red")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Total number of customers in the trial period for the trial store is significantly
higher than the control store for two out of three months, which indicates a
positive trial effect.
## Conclusion
Good work! We've found control stores 233, 155, 237 for trial stores 77, 86 and 88
respectively.
The results for trial stores 77 and 88 during the trial period show a significant
difference in at least two of the three trial months but this is not the case for
trial store 86. We can check with the client if the implementation of the trial was
different in trial store 86 but overall, the trial shows a significant increase in
sales. Now that we have finished our analysis, we can prepare our presentation to
the Category Manager.